PhD Thesis

### 2016-10-24 pldi14/sec-practice.tex
Integrate PLDI14 paper
Minor fixes to PLDI14

\input{fig-case-study-pseudocode}

@@

\input{sec-performance}

@@

\input{fig-primitives}

\pg{Why does 'target language' show up all of a sudden? Shouldn't
  we say that we generate Scala code for execution
  \textbf{elsewhere}? Right now, we say it in next section. Confusing.}
%
\Cref{fig:primitives} shows a simplified
Scala implementation of the primitives
used in \cref{fig:case-study-pseudocode}.
As bag primitives, we provide constructors and a fold operation,
following \citet{GlucheGrust97Incr}. The constructors for bags
are $\Empty$ (constructing the empty bag), $\SINGLETON$
(constructing a bag with one element), $\MERGE$ (constructing the
merge of two bags) and $\NEGATE$ ($\Negate{b}$ constructs a bag
with the same elements as $b$ but negated multiplicities); all but $\SINGLETON$ represent
abelian group operations.
%
Unlike for usual ADT constructors, the same bag can be
constructed in different ways, which are equivalent by the equations defining abelian groups;
for instance, since $\MERGE$ is commutative, $\Merge{x}{y} = \Merge{y}{x}$.
%
Folding on a bag will represent the bag through constructors in
an arbitrary way, and then replace constructors with arguments;
to ensure a well-defined result, the arguments of fold should
respect the same equations, that is, they should form an abelian group;
for instance, the binary operator should be commutative.
%
Hence, the fold operator $\FOLDBAG$ can be defined to take a
function (corresponding to $\SINGLETON$) and an abelian group
(for the other constructors). $\FOLDBAG$ is then defined by equations:
%
\begin{alignat*}{2}
&  \HasType{\FOLDBAG}{\mathbf{Group}\; \Gt \to (\Gs \to \Gt) \to &&\Bag[\Gs] \to \Gt}\\
&  \App{\FoldBag {g @ (\_, \boxplus, \boxminus, e)}{f}}{\Empty}
    &&= e \displaybreak[0]\\
&  \App{\FoldBag {g @ (\_, \boxplus, \boxminus, e)}{f}}{\Merge* {b_1} {b_2}}
   &&= \App{\FoldBag{g}{f}}{b_1} \displaybreak[0]\\
&&& \boxplus
  \App{\FoldBag{g}{f}}{b_1} \displaybreak[0]\\
&  \App{\FoldBag {g @ (\_, \boxplus, \boxminus, e)}{f}}{\Negate* b}
   && = \boxminus \; \App*{\FoldBag{g}{f}}{b}\displaybreak[0]\\
&  \App{\FoldBag {g @ (\_, \boxplus, \boxminus, e)}{f}}{\Singleton*{v}}
   &&= \App{f}{v}
\end{alignat*}
%
If $g$ is a group,
these equations specify $\App{\FOLDBAG} g$ precisely~\citep{GlucheGrust97Incr}.
%
Moreover, the first three equations mean that $\FoldBag{g}{f}$ is \emph{abelian group
  homomorphism} between the abelian group on bags and the group
$g$ (because those equations coincide with the definition).
%
\Cref{fig:primitives} shows an implementation of $\FOLDBAG$ as
specified above.
Moreover, all functions which deconstruct a bag can be expressed in
terms of $\FOLDBAG$ with suitable arguments.
%
For instance, we can sum the elements of a bag of integers with
$\FoldBag{\Term{gZ}}{\Lam*{x}{x}}$, where
$\Term{gZ}$ is the abelian group on integers defined in
\cref{ssec:change-structures}.
Users of $\FOLDBAG$ can define different abelian groups to specify
different operations (for instance, to multiply floating-point numbers).

@@

To define $\Term{GroupChange}$, we need a suitable erased change
structure on $\Gt$, such that $\UPDATE$ will be equivalent to
$\boxplus$. Since there might be multiple groups on $\Gt$, we
\emph{allow the changes to specify a group}, and have
$\UPDATE$ delegate to $\boxplus$:
\begin{align*}
& \Change{\Gt} = \Term{Replace}\; \Gt \mid \Term{GroupChange} \Abelian*{\Gt} \Gt \\
& \Update{v}{(\Term{Replace} \; u)} = u\\
& \Update{v}{(\Term{GroupChange} \; (\bullet, \Term{inverse}, \Term{zero})\; dv)} = v \bullet dv\\
& \Diff{v}{u} = \Term{Replace}\; v
\end{align*}
That is, a change between two values is either simply the new
value (which replaces the old one, triggering recomputation),
or their difference (computed with abelian group
operations, like in the changes structures for groups from
\cref{ssec:change-structures}. The operator $\DIFF$
does not know which group to use, so it does not take advantage
of the group structure.
However, $\FOLDBAG$ is now able to generate a group change.
% Rewrite:
%derivatives of primitives like $\FOLDBAG$ can
%use the group structure they have available when producing changes.
%
\pg{Clarify about user-defined groups.}

@@

\begin{figure}
\includegraphics[keepaspectratio,width=8.5cm]{HistogramGenerated-new.pdf}
\caption{Performance results in log-log scale, with input size on
  the x-axis and runtime in ms on the y-axis. Confidence
  intervals are shown by the whiskers; most whiskers are
  too small to be visible.}
\label{fig:graph}
\end{figure}

>>>>

\input{pldi14/fig-case-study-pseudocode}

@@

\input{pldi14/sec-performance}

@@

\input{pldi14/fig-primitives}

\pg{Why does 'target language' show up all of a sudden? Shouldn't
  we say that we generate Scala code for execution
  \textbf{elsewhere}? Right now, we say it in next section. Confusing.}
%
\Cref{fig:primitives} shows a simplified
Scala implementation of the primitives
used in \cref{fig:case-study-pseudocode}.
As bag primitives, we provide constructors and a fold operation,
following \citet{GlucheGrust97Incr}. The constructors for bags
are $\Empty$ (constructing the empty bag), $\SINGLETON$
(constructing a bag with one element), $\MERGE$ (constructing the
merge of two bags) and $\NEGATE$ ($\Negate{b}$ constructs a bag
with the same elements as $b$ but negated multiplicities); all but $\SINGLETON$ represent
abelian group operations.
%
Unlike for usual ADT constructors, the same bag can be
constructed in different ways, which are equivalent by the equations defining abelian groups;
for instance, since $\MERGE$ is commutative, $\Merge{x}{y} = \Merge{y}{x}$.
%
Folding on a bag will represent the bag through constructors in
an arbitrary way, and then replace constructors with arguments;
to ensure a well-defined result, the arguments of fold should
respect the same equations, that is, they should form an abelian group;
for instance, the binary operator should be commutative.
%
Hence, the fold operator $\FOLDBAG$ can be defined to take a
function (corresponding to $\SINGLETON$) and an abelian group
(for the other constructors). $\FOLDBAG$ is then defined by equations:
%
\begin{alignat*}{2}
&  \HasType{\FOLDBAG}{\mathbf{Group}\; \Gt \to (\Gs \to \Gt) \to &&\Bag[\Gs] \to \Gt}\\
&  \App{\FoldBag {g @ (\_, \boxplus, \boxminus, e)}{f}}{\Empty}
    &&= e \displaybreak[0]\\
&  \App{\FoldBag {g @ (\_, \boxplus, \boxminus, e)}{f}}{\Merge* {b_1} {b_2}}
   &&= \App{\FoldBag{g}{f}}{b_1} \displaybreak[0]\\
&&& \boxplus
  \App{\FoldBag{g}{f}}{b_1} \displaybreak[0]\\
&  \App{\FoldBag {g @ (\_, \boxplus, \boxminus, e)}{f}}{\Negate* b}
   && = \boxminus \; \App*{\FoldBag{g}{f}}{b}\displaybreak[0]\\
&  \App{\FoldBag {g @ (\_, \boxplus, \boxminus, e)}{f}}{\Singleton*{v}}
   &&= \App{f}{v}
\end{alignat*}
%
If $g$ is a group,
these equations specify $\App{\FOLDBAG} g$ precisely~\citep{GlucheGrust97Incr}.
%
Moreover, the first three equations mean that $\FoldBag{g}{f}$ is the \emph{abelian group
  homomorphism} between the abelian group on bags and the group
$g$ (because those equations coincide with the definition).
%
\Cref{fig:primitives} shows an implementation of $\FOLDBAG$ as
specified above.
Moreover, all functions which deconstruct a bag can be expressed in
terms of $\FOLDBAG$ with suitable arguments.
%
For instance, we can sum the elements of a bag of integers with
$\FoldBag{\Term{gZ}}{\Lam*{x}{x}}$, where
$\Term{gZ}$ is the abelian group on integers defined in
\cref{ssec:change-structures}.
Users of $\FOLDBAG$ can define different abelian groups to specify
different operations (for instance, to multiply floating-point numbers).

@@

To define $\Term{GroupChange}$, we need a suitable erased change
structure on $\Gt$, such that $\UPDATE$ will be equivalent to
$\boxplus$. Since there might be multiple groups on $\Gt$, we
\emph{allow the changes to specify a group}, and have
$\UPDATE$ delegate to $\boxplus$ (which we extract by pattern-matching on the group):
\begin{align*}
& \Change{\Gt} = \Term{Replace}\; \Gt \mid \Term{GroupChange} \Abelian*{\Gt} \Gt \\
& \Update{v}{(\Term{Replace} \; u)} = u\\
& \Update{v}{(\Term{GroupChange} \; (\boxplus, \Term{inverse}, \Term{zero})\; dv)} = v \boxplus dv\\
& \Diff{v}{u} = \Term{Replace}\; v
\end{align*}
That is, a change between two values is either simply the new
value (which replaces the old one, triggering recomputation),
or their difference (computed with abelian group
operations, like in the changes structures for groups from
\cref{ssec:change-structures}. The operator $\DIFF$
does not know which group to use, so it does not take advantage
of the group structure.
However, $\FOLDBAG$ is now able to generate a group change.
% Rewrite:
%derivatives of primitives like $\FOLDBAG$ can
%use the group structure they have available when producing changes.
%
\pg{Clarify about user-defined groups.}

@@

\begin{figure}
\includegraphics[keepaspectratio,width=8.5cm]{pldi14/HistogramGenerated-new.pdf}
\caption{Performance results in log-log scale, with input size on
  the x-axis and runtime in ms on the y-axis. Confidence
  intervals are shown by the whiskers; most whiskers are
  too small to be visible.}
\label{fig:graph}
\end{figure}
